---
title: "Exercise 1"
author: "Carilli"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
library(readr)
library(janitor)
library(broom)
library(magrittr)
library(kableExtra)
library(lm.beta)
library(jtools)
```

```{r data}
instability <- read_csv("Instability_PS data.csv")
```

### (a)

```{r part_a}
instability %$% 
  lm(instab ~ democracy) %>% 
  lm.beta()

## or

instability %>% 
  mutate(instab = scale(instab),
         democracy = scale(democracy)) %$% 
  lm(instab ~ -1 + democracy) %>% 
  tidy()

## or

instability %$%
  lm(instab ~ democracy) %>% 
  summ(scale = TRUE, digits = 3)
```

### (b)

```{r part_b}
instability %>% 
  mutate(instab = scale(instab),
         democracy = scale(democracy),
         gdplag = scale(gdplag)) %$% 
  lm(instab ~ -1 + democracy + gdplag) %>% 
  tidy()
```

### (c)

```{r part_C}
instability %>% 
  mutate(instab = scale(instab),
         democracy = scale(democracy),
         gdplag = scale(log(gdplag))) %$% 
  lm(instab ~ -1 + democracy + gdplag) %>% 
  tidy()
```

### (d)

```{r part_d}
instability %>% 
  mutate(instab = scale(instab),
         democracy = scale(democracy),
         gdplag = scale(log(gdplag))) %$% 
  lm(instab ~ -1 + democracy + gdplag + coldwar) %>% 
  tidy()
```

### (e)

```{r part_e}
instability %$% 
  lm(instab ~ democracy + I(democracy^2) + gdplag + coldwar) %>% 
  tidy()

beta <- 
instability %$% 
  lm(instab ~  democracy + I(democracy^2) + gdplag + coldwar) %>% 
  tidy() %>% 
  pull(estimate)
```

Our model is $\text{instab} = \hat{\beta_1}\text{democracy} + \hat{\beta_2}\text{democracy}^2 + \hat{\beta_3}\text{gdplag} + \hat{\beta_4}\text{coldwar}$, to find the point after which increases in democracy lead to instability find the parital derivitive of instability with respect to democracy and set it equal to 0.

\begin{aligned}
\frac{\partial\text{ instab}}{\partial\text{ democracy}} &= \hat{\beta_1} + 2\hat{\beta_2}\text{ democracy} = 0\\\\
2\hat{\beta_2}\text{ democracy} &= -\hat{\beta_1} \\\\
\text{democracy} &= - \frac{\hat{\beta_1}}{2\hat{\beta_2}} \\\\
& = `r -beta[2]/(2*beta[3])`
\end{aligned}

```{r}
tibble(
democracy = 1:100,
instability = beta[2] * (1:100) + beta[3] * (1:100)^2
) %>% 
  ggplot(aes(x = democracy, y = instability)) + 
  geom_line() +  
  geom_vline(xintercept = -beta[2]/(2*beta[3]), linetype = "dotted") + 
  labs(x = "Democracy Score", y = "Predicted Instability")
```

